Busch, Khakhalin. Intrinsic temporal tuning

### 2019-06-03 reviews and responses.md

### Minor

> Page 5, line 45/46: “As one possible explanation, auditory and mechanosensory inputs may have
different cellular or subcellular 46 targets in the tectum (Bollmann and Engert, 2009).”
This Bollmann and Engert reference does not address mechanosensory and auditory inputs. This paper
shows that the RGC inputs form a topographic map along the proximal-distal axis of individual tectal
neuron dendrites. This reference should be replaced with an appropriate reference.

> Also, how can having different subcellular targets could explain the results? This concept needs to be
clarified – an example would be helpful.

We replaced this sentence with the following one: "This was not necessarily surprising, as mechanosensory and visual inputs have different cellular targets in the tectum (Pratt and Aizenman, 2009; Felch et al., 2016; Truszkowski et al., 2017), potentially leading to different recruitment of tectal excitatory and inhibitory circuits. Acoustic stimuli may also be inherently weaker than visual stimuli in triggering plasticity effects in tectal neurons, as they arrive at different compartments within the dendritic tree (Hiramoto and Cline, 2009; Deeg et al., 2009), which may shape their influence on neuronal plasticity (Richards and Lillicrap, 2019)."

Here, the actual rationale for the second explanation (a difference in subcellular targets) is that inputs to different parts of the dentritic tree can trigger action potentials of different shape, and differentially affect calcium dynamics in the neuron, which, together with non-linear processing in dendrites may make different types of inputs unequal in terms of their effects on neuronal activity, and neuronal tuning, respectively. Two extreme examples of this effect would be the impact of climbing fibers on the activity of Purkinje cells in the cerebellum, and the impact of inputs to the apical dendrites of pyramidal cells in the cortex. We feel, however, that this topic is really deep, yet somewhat tangential to the main message of the paper, so we tried to limit our exposition to it by short sentence and a good recent reference.

@@

> Page 13 line 46: The authors use the term “distributed sensory network” . Is a “distributed” network a
special type of sensory network?

Technically, even two neurons form a network (and a true mathematician would probably argue that even one neuron forms a network!), yet the tectal network is a relatively large one, with broad recurrent excitation, and in which different parts of the sensory input are processed in parallel. To subtly remind the reader about that, we used a word "distributed", as in "distributed computations", as opposed to more local and covergent networks, such as, for example, the Mauthner cell network in the hindbrain. This word is often used in writing (a Google Scholar search for `"distributed network" brain` returns 27000 hits), but admittedly it is not a precisely defined term, but rather a useful qualifier. If you think that this word is not appropriate here, please let us know, and we will remove it.

@@

> Page 3, line 26,27,28: “In the tadpole tectum, looming stimuli are known to elicit stronger responses
compared to instantaneous inversions, yet the suppression yielded by looming stimuli was weaker”.
The logic of this sentence is not clear. Why exactly would the authors not expect that the looming
stimuli that is known to elicit stronger (synaptic?) responses would also cause less suppression (of action
potentials?) than does flashes?

This was unexpected as intrinsic plasticity is typically described as a homeostatic mechanism, and thus a stronger stimulus could, in principle, cause stronger suppression of spiking. To make this point more clear, we now changed this sentence:

"This was likewise unexpected, as compared to flashes, looming stimuli are known to elicit stronger tectal responses (Khakhalin et al., 2014; Khakhalin, 2019), yet their effect on neuronal excitability was weaker, seemingly in contradiction with the principle of homeostatic plasticity (Pratt and Aizenman, 2007; Turrigiano, 2007)."

@@

### Major

> Page 5, line 2,3,4 : But do the authors actually know how/if their conditioning stimuli is altering the
pattern/strength of synaptic input? If so, this data should be shown. If not, it should be made clear that
this is an assumption/ theoretical prediction.

In the originally submitted pdf, lines 2 to 5 read:

"We then examined whether different types of sensory activation would differentially reshape  
temporal intrinsic tuning in tectal neurons. As changes in intrinsic properties seemed  
homeostatic (increased activation led to reduced spiking), by the same logic, one could
expect shorter stimuli (flashes) to selectively suppress responses to shorter synaptic inputs."

Here we are not implying that our conditional stimuli altered the pattern of synaptic inputs; rather, we refer to the fact that _during_ the stimulation, shorter stimuli (such as instantaneous flashes, which technically makes them the shortest stimuli possible) would cause faster visual inputs to the tectum, compared to longer, continuous transitions (second-long looms). Within this study, we did not measure actual synaptic responses in the tectum during conditioning, but from previous studies, we know that responses to flashes are indeed much shorter than responses to looming stimuli (Khakhalin 2014, Khakhalin 2019). We refer to these studies in the beginning of the "Results" section, and then again in the "Methods".

For page 5, we now changed the wording, to make this logic clearer.

"As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (instantaneous flashes; Khakhalin et al. 2014) would selectively suppress spiking in response to faster synaptic inputs."

@@

> Page 5, line 5: “As changes in intrinsic properties seemed homeostatic (increased activation led to
reduced spiking), by the same logic, one could expect shorter stimuli (flashes) to selectively suppress
responses (action potential firing or synaptic responses or both?) to shorter synaptic inputs”.
What intrinsic properties were measured? It appears that only number of spikes was counted. The
authors report that long-term increase in activation (by visual stimuli) decreased spiking – but was that
due to a decrease in intrinsic excitability or because the stimuli decreased synaptic strengths – which in
turn led to the observed decrease in number of spikes fired? This is critical to define.
Second, do the authors mean to say that shorter stimuli would actually (somehow) suppress responses?
How would that work?

Our measurements of spiking output relied on dynamic clamp experiments, in which we used simulated synaptic inputs, and not real synaptic stimulation. It means that any changes in spiking output we report in this paper are necessarily coming from the change in intrinsic properties. While a change in the strength of synaptic inputs  in response to sensory stimulation is possible, and indeed was observed and investigated in later sections of this paper, it would not have affected the number of spikes we reported in figures 1 and 2, as they were recorded in dynamic clamp experiments.

With the new wording of this sentence, we tried to make this point more clear:

"As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (instantaneous flashes; Khakhalin et al. 2014) would selectively suppress spiking in response to faster synaptic inputs."

@@

> Page 5 line 15: Does the word “preferred” (in quotes) mean that the cells fired more action potentials?
Maybe refrain from personification.

We hope that the word "prefer" is intuitive, and it has a benefit of being way shorter than "responded with more spikes", or "produced more action potentials", which hopefully makes reading easier. Realistically, we hope that most readers of this article will not be at risk of truly personifying neurons, so while we take care to never personify animals, we are intentionally a bit more relaxed when writing about cells, channels, receptors etc. Still, to make this sentence better, we removed the quotes around "preferred", to make it more obvious that it is not an antropomorphic allegory, but just a normal word use. Overall, we use the word "prefer" 6 times in the paper, always in the same way: when comparing spiking responses to different stimuli, to describe the stimulus that produced sronger outputs in a given set of neurons.

@@

These are the same stimuli, and we changes these two sentences to better reflect that:

"To test this question, we exposed tadpoles to four hours of behaviorally salient, startle-inducing “click” sounds, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to these sounds..."

> - Tadpoles are exposed to 4 hours of click sounds and then intrinsic tuning was measured. No difference
in spiking was observed. This was interpreted to mean that acoustic stimuli did “not strongly activate
tectal circuits during conditioning”. How can the authors know this? Instead, why not determine this
directly by recording click-evoked responses from tectal neurons? Or, if tectal cell responses to auditory
stimuli have been described in previous work, this should be included here.

Unfortunately, electrophysiological recording of tectal reponses to sounds in tadpoles does not seem to be technically feasible, not just in our lab (where we don't even have the technical capability to perform in-vivo experiments), but in most labs. From the technical point of view, the problem amounts to two limitations: one, to record from a tadpole, one has to paralyze it with a nicotinic receptor anatagonist, which seems to disrupt hearing. Second, even if this limitation could be overcome, the sounds have to be strong enough to stimulate the inner ear, but weak enough not to mechanically disrupt the recording. Conceivably, both challenges may circumvented with the use of high speed calcum imaging in freely moving tadpoles, but to our knowledge no lab has attempted these recordings yet.

@@

> Page 6 line 24: Remove the phrase “are not easily interpretable”. Instead, please interpret your data
clearly.

We removed this phrase.

> Page 9: The authors report that the synaptic data show that “prolonged sensory activation with short,
frequent stimuli reshaped synaptic transmission in the tectum, making it faster through selective
potentiation of visual inputs from the eye”.
How do the authors know the effect is selective for RGC inputs? It appears that they only measured
RGC-evoked responses. Auditory/mechanosensory -evoked responses were not reported.

Stimulation in the optic chiam evokes two types of synaptic responses in tectal cells: direct short-latency responses, and delayed recurrent inputs from other cells in the tectum. By looking at average currents early and later after the chiasm shock, we can independently assess the contribution of direct and recurrent inputs. In the paper, in a paragraph before the sentence in question, we report that after stimulation, direct current increased in amplitude, while recurrent currents remained unchanged. This is what we call selective potentiation of direct inputs, as opposed to recurrent inputs. We do not claim that mechanosensory inputs did not change, as we have not recorded auditory responses in this study (recording of auditory responses is considerably more difficult; see Felch 2016).

To make sure that we are not misunderstood by the reader, we replaced the original sentence with the following two sentences: "This shows that prolonged sensory activation with short, frequent stimuli reshaped synaptic transmission in the tectum, making it faster through selective potentiation of visual inputs from the eye, compared to recurrent inputs within the tectum, that remained unchanged. In this study, we did not probe changes in acoustic and mechanosensory projections from the hindbrain."

@@

> Page 9 line 9: “cells exposed to stronger stimuli preferred shorter synaptic inputs in dynamic clamp…”
Here, (where “preferred” is not in quotes) does “preferred” mean to fire more action potentials?

As described above, we now don't use quotation marks for any basic words, so "preferred" is used without quotation marks, and consistently means "responded preferably", or "produced stronger responses".

@@

> Page 10 (last paragraph): Were action potential thresholds measured? Were there any differences in
this parameter across groups?

We did not measure action potential thresholds, as we know from the previous study (Ciarleglio Khakhalin 2015) that threshold measurements in tadpole tectal neurons are highly sensitive to the current that is used to evoke the spikes, due to rapid Na channels inactivation (Azouz 2000; Fontaine 2014).

@@

> Page 10 line 8: The “correction for position” is difficult to understand here and in Methods. Why was a
linear regression used – and how is it determined to generate an accurate “correction”? How do the
data look without this correction?

It was a mental typo: we should have used the standard term "adjusted for position"; thank you for noticing that! We performed a standard adjustment for a covariate, as used in ANCOVA and other types of general linear model analysis. We used a linear model, as it is a "default" approach when the exact relation between two variables is unknown, and the amount of noise is high enough to make the difference between linear and curvilinear regressions negligible (both cases apply here).

There is no way for us to determine whether our adjustment is "accurate", as there is no clear way to make more caudal (and thus developmentally younger) cells in the tectum older. This situation is not unique however, as it arises in any field that uses linear models, from ecology to social sciences. When running a study in a human population, if a variable of interest depends on age, one would typically adjust for age by including age as a covariate in the linear model. Here we perform the same standard procedure. After rereading the manuscript, we realized that we used terms "adjustment", "correction", and, in one case, "compensation" interchangeably, which obviously led to confusion. We now consistently use the term "adjustment" throughout the manuscript.

In the paper, we do not report how the data would have looked without an adjustment for covariates, as it would not be statistically sound to run a linear model with a known covariate without including this covariate in the analysis. For this response, we ran some of the analyses without a covariate, and did not find qualitative differences, although numerical values of "uncorrected" d, F and p values were of course different from those reported in the paper. For example, when the measure of "spikness" (one of the three tuning parameters) is compared across groups, a correct analysis with adjustment for position (`smean~rostral+medial+group`) produces a p-value of 0.01. Without adjustment (`smean~group`) we get a p-value of 0.03. For temporal tuning, correct analysis (`sbend~rostral+medial+group`) gives p=0.030, and if covariate is left unadjusted (`sbend~group`) we get p=0.029. For amplitude tuning, we get p=0.0021 for correct analysis, and p=0.0024 for unadjusted analysis. For the analyses in which temporal tuning and ampliltude coefficients were explained through intrinsic parameters, optimal models were not different between adjusted and unadjusted data, but all values in the table were slightly different. If you would like us to numerically re-run any other part of the analysis without a position adjustment, please let us know which part you would like to see. Also, all scripts that produce technical checks, visualizations, and analyses related to this adjustment are available for inspection at:

https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/dynamic_other_vars.R

@@

> Page 13, line 7: What does “intrinsic spikiness” mean? Is this another term for intrinsic excitability, or is
this a different parameter? In fact, throughout the manuscript, there are many terms including “intrinsic
excitability, spikiness, number of spikes, and intrinsic spikiness that probably mean the same thing – or
maybe not. If these terms all mean the same thing, then one term should be chosen and used
throughout. If they all refer to different entities, then each one should be clearly defined.

Thank you for making us get more consistent here! In this updated version, we consistently use "spikiness" as a name for the variable that is for the first time introduced in Figure 2C. When writing about the average number of spikes, we now just call it the "average number of spikes", and not "spikiness". As the only exception to this rule, we opted for the word "spikiness" in the first paragraph of Introduction, to prime the reader for the use of this word later in the text, even though strictly speaking it is a sentence about intrinsic excitability. The expression "intrinsic spikiness" is no longer used. The expression "intrinsic excitability" is now used only as a name for the underlying property of the neuron, that may be probed with different alternative protocols and measures. In this study, we mostly assess intrinsic excitability using a more narrowly defined "spikiness" variable, but towards the end of the paper, we also compare it to a less informative measure of the number of spikes observed in current clamp experiments. We use "intrinsic excitability" when we make inferences about the actual properties of the neuron, and use "spikiness" when we talk about the actual numerical value calculated to assess and quantify intrinsic excitability.

@@

> Page 15 line 34: The average membrane resistance (330 MOhm) seems a bit low for these small immature
tectal neurons. Were the neurons healthy? What was the average resting membrane potential? Were
there any set criteria for Rm, Vm, or access resistance?

Thank you for catching this error! We had an order of magnitude typo here, as the average membrane resistance was in fact 3.2 GOhm. This value is a bit higher than reported in earlier studies (1.2±0.8 GOhm in Ciarleglio Khakhalin 2015; about 1.1 in Khakhalin 2011), but within what can be reasonably expected for healthy neurons of this size. To make sure that this number is right, for the purpose of this response, we also verified it by reanalyzing our data offline, and obtained a similar estimation for the average membrane resistance (2.9±1.4 GOhm).

We have several reasons to believe that the cells included in this study were healthy. First, the access resistances obtained were good (30±14 MOhm, which is similar to benchmarks of 50±23 in Ciarleglio Khakhalin 2015, and about 20-30 MOhm in Khakhalin 2011), indicating that higher membrane resistance is not likely to be due to poor access. Second, our neurons reliably produced spikes in current clamp mode, and active currents in response to membrane depolarization in voltage clamp mode. Moreover, our quantification of these active conductances yielded numbers similar to those reported before (Ciarleglio Khakhalin 2015). The quality of voltage clamp, as judged by the absence of escaping active currents in IV protocols, was good in the majority of neurons (82% of all neurons recorded). Finally, a few cells that had their membrane resistance change drastically during the recording, were discarded at the data collection stage, and not included in this study.

We did not measure the resting membrane potential directly, but measured the holding current necessary to bring the cell to -60 mV in voltage clamp mode, and the mean value of this I_hold value was -8.2±4.8 pA, which is similar (most importnatly, not higher) than what was previously reported in the literature (-19.8±16.6 in Ciearleglio Khakhalin 2015).

Together, these arguments make us believe that the quality of our patches was good enough to draw the conclusions reported in the study. Note that all data for all cells is available here:

https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/Data/data_mainInput.txt

In this full dataset, groups "Naive", "SlowF", and "SlowC" should be ignored, as they were not included in the study. Note also that if you open this file on github, the columns may look misplaced, as github does not render tab-separated files correctly, so if you want to inspect the data, we recommend that you download it and open it in a spreadsheet.

As the mistake your noticed (thank you again for being so thorough!) was a typo in the text, and not a mistake in the data, it does not affect any calculations reported in this paper.

@@

> Figure 1 D and E: The average number of spikes fired as a function of conductance duration and all at 3
different peak conductances. Are there any trends at the single cell level? For example, if a given
control neuron fires relatively few spikes in response to 500 ms duration, does it also tend to fire
relatively fewer spikes with 1000ms duration? Or maybe if a given neuron fires a relatively high number
of spikes with 500ms then it will typically fire less at 1000ms duration? Although the individual
responses are shown in 1D, it is impossible to follow the number of spikes fired by any individual neuron
in response to the different conductance durations.

At least for high-spiking neurons, the answer to this question can be inferred from Figure 1D: those curves that are placed high on the graph for duration of 500 ms, also tend to remain high for 1000 ms, indicating that high-spiking neurons tend to be high-spiking across the range of durations. Numerically, the correlation coefficient between the average number of spikes in response to 500 and 1000 ms, across all neurons, is r=0.94. This is to be expected, as intrinsic excitability of neurons (whether the neuron is spiky or not) is a much stronger feature than the temporal tuning. (Temporal tuning in the tectum was not even described until this paper, while the variation in intrinsic excitability in the tectum is well known for more than 20 years).

@@

> Figure 4 D, E and F: Is “spikiness” synonymous with “N Spikes”? If these terms do not mean the same
thing, they should be defined. If they do mean the same thing, then one term should be chosen and
used consistently throughout.

As described above, "spikiness" is now consistently used across the manuscript as the name of one of three variables used to quantify intrinsic tuning of tectal neurons (alongside with temporal tuning and amplitude tuning).

@@

We now made the introduction clearer, and better organized, going from the big problem, to specific questions. We moved the presentation of the questions for the study to the very beginning of the second paragraph, and provided a rationale for these questions in the first paragraph.

> Results and Discussion:

> - The rationale for the use of dynamic clamp and the parameters chosen are not clear.

### 2019-06-03 main.tex
Tex updates: routine improvements

After conditioning, we excised the brain, performed whole cell patch clamp on tectal neurons (Figure 1A), and counted spikes produced in response to dynamically simulated synaptic conductances (Figure 1B). We used conductances of 4 different durations (100, 200, 500, and 1000 ms), and 3 different amplitudes (peak conductances of 0.2, 0.5, and 1.0 nS), matching the range of synaptic currents observed in tectal circuits \textit{in vivo} (\citealt{xu2011,khakhalin2014,ciarleglio2015}; see Methods). Based on prior research \citep{aizenman2003,ciarleglio2015}, we expected neurons exposed to instantaneous checkerboard inversions (flashes) to become more excitable, but surprisingly, they were on average less spiky, and generated 0.4$\pm$0.4 spikes, across all types of dynamic clamp inputs, compared to 0.9$\pm$1.0 spikes in control (Figure 1D, left column; F(1,677)=30.4, p=5e-8, n cells=28, 29). (Here and below, F-values are reported for a multivariate fixed effects analysis of variance with selected interactions, where cell id is used as a repeated measures factor; see Methods for a detailed description). Neurons exposed to looming transitions also spiked less than control neurons (0.6$\pm$0.4 spikes; F(1,629)=9.6; p=0.002; n=28, 25), but more than those exposed to instantaneous "flashes" (F(1,641)=3.7; p=4e-6; n=29, 25). This was likewise unexpected, as compared to flashes, looming stimuli are known to elicit stronger tectal responses \citep{khakhalin2014,khakhalin2019graph}, yet their effect on neuronal excitability was weaker, seemingly violating the principle of homeostatic plasticity \citep{pratt2007,turrigiano2007}.

@@

We then examined whether different types of sensory activation would differently reshape temporal intrinsic tuning in tectal neurons. As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (flashes; \citealt{khakhalin2014}) would selectively suppress spiking in response to shorter synaptic inputs. Alternatively, shorter stimuli could make neurons better adjusted to working with short bursts of activation \citep{stemmler1999information}, as  described for synaptic processing \citep{aizenman2007} and recurrent activity in the tectum \citep{pratt2007,shen2011}.

We found that compared to control, neurons exposed to flashes and looming stimuli had different response curves when tested across input conductances of different lengths (F(1, 677)=25.1; p=7e-7; and F(1,629)=12.0; p=6e-4 respectively). In control neurons, longer inputs (500 and 1000 ms) typically evoked stronger spiking than shorter inputs (100 and 200 ms), but neurons from stimulated animals had flatter tuning curves, with a plateau, or even a decrease in spike output for longer conductance injections (Figure 1E). In other words, while control neurons preferred longer inputs, stimulated neurons developed a preference for shorter synaptic inputs, and this change was more pronounced in neurons exposed to flashes than in those exposed to looming stimuli (F(1,641)=7.5; p=0.006). This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, follow two different kinds of logic. The overall excitability is homeostatic, as neurons became less excitable in response to stronger stimulation. The temporal retuning however can be better described as “adaptive”, as neurons exposed to shorter stimuli (flashes) became relatively \textit{more} responsive to shorter stimuli, and less responsive to longer stimuli, making them more equipped to process faster, more synchronous patterns of activation \citep{stemmler1999information,fontaine2014threshold}.

@@

To test this question, we exposed tadpoles to four hours of behaviorally salient, startle-inducing “click” sounds \citep{james2015,truszkowski2017}, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to these sounds (group “Sound”) did not lead to significant changes in either average number of spikes (0.8$\pm$0.7 spikes; F(1,689)=1.7; p=0.2; n=28, 30), amplitude transfer function (F(1,689)=2.0; p=0.2), or temporal tuning curve (F(1,689)=2.0; p=0.2). This may suggest that acoustic stimuli did not activate tectal circuits strongly enough during conditioning, despite being more behaviorally salient (at the onset of stimulation, acoustic clicks evoked startle responses in about 50-80\% of cases, compared to 5-10\% for checkerboard inversions \citep{james2015,truszkowski2017}). This was not necessarily surprising, as auditory and mechanosensory inputs are known to have different cellular targets in the tectum \citep{pratt2009trigeminal,felch2016,truszkowski2017}, and thus can recruit tectal excitatory and inhibitory networks to a different extent. Alternatively, acoustic stimuli may be inherently weaker in triggering plasticity effects in tectal neurons, as compared to visual inputs, they arrive at different subcellular compartments within the dendritic tree \citep{hiramoto2009,deeg2009}, which can affect their influence on neuronal plasticity \citep{richards2019dendritic}.

@@

To better match and compare temporal properties of synaptic inputs to those of intrinsic tuning, we calculated average "synaptic current duration" for each cell as a temporal “center of mass” of currents within the first 700 ms after optic chiasm stimulation (see Methods). Neurons with different contribution of early and late synaptic responses naturally had different synaptic current duration: cells with strong monosynaptic inputs had shorter currents, while polysynaptic activity made synaptic currents longer (Figure 3C; p=2e-16, r=$-$0.78, n=168). The synaptic current duration was different across treatment groups (Figure 3D; F(5,163)=6.3, p=2e-5). Cells in Flash, Sync, and Async groups all received shorter synaptic inputs than Control cells (Tukey p$<$0.05, mean duration of 267$\pm$36, 253$\pm$39, 268$\pm$24, and 304$\pm$40 ms, respectively). This shows that prolonged sensory activation with short, frequent stimuli reshaped synaptic transmission in the tectum, making it faster through selective potentiation of visual inputs from the eye, compared to recurrent inputs within the tectum that remained unchanged. In this study, we did not perform a similar analysis for acoustic projection from the hindbrain.

@@

As our technical resources were rather limited, in this study we don't explicitly address the mechanisms of this newly discovered intrinsic temporal tuning. We can, however, offer two working hypotheses that may explain these results. As tadpole tectal cells don't express "true" resonance currents, such as h-currents \citep{ciarleglio2015}, most temporal tuning effects we observed seem to be due to differences in ionic current inactivation in different cells \citep{azouz2000threshold,fontaine2014threshold}. One obvious way of tuning ionic channel inactivation would be for every cell to modulate sodium or transient potassium currents via channel phosphorylation, or by changing the expression of different channel variants, to shift their inactivation dynamics \citep{frank2003nachannels,goldwyn2018a_current}. To test this hypothesis, one would need to induce changes in temporal tuning similarly to how we did in this paper, pharmacologically isolate different ionic currents, and directly measure their inactivation dynamics \citep{zbili2019axonnav}.

@@

Our findings lead to several verifiable predictions. As rapid inactivation of spiking in tectal neurons plays a role in collision detection \citep{khakhalin2014,jang2016}, a change in temporal tuning should affect collision detection dynamics, which can be verified experimentally. More specifically, we predict that a visual stimulation that retunes neurons to faster stimuli (Flash) would make tadpoles selectively less responsive to slow collisions, and increase the latency of collision avoidance due to non-linear dynamics of retinal activation for realistic looming stimuli that are slow in the beginning and fast towards the end \citep{jang2016}. These changes in intrinsic temporal tuning would also reshape the connectivity of tectal networks, as fast-inactivating cells would not support short recurrent loops within the network, thus promoting long-ranged polysynaptic connectivity \citep{fiete2010stdp,clopath2010stdp}. Finally, based on the multisensory phenomena reported in this paper, we predict that even though multisensory stimulation tends to increase tectal responses in vitro \citep{felch2016,truszkowski2017}, it is be likely to reduce peak activation in vivo.

>>>>

After conditioning, we excised the brain, performed whole cell patch clamp on tectal neurons (Figure 1A), and counted spikes produced in response to dynamically simulated synaptic conductances (Figure 1B). We used conductances of 4 different durations (100, 200, 500, and 1000 ms), and 3 different amplitudes (peak conductances of 0.2, 0.5, and 1.0 nS), matching the range of synaptic currents observed in tectal circuits \textit{in vivo} (\citealt{xu2011,khakhalin2014,ciarleglio2015}; see Methods). Based on prior research \citep{aizenman2003,ciarleglio2015}, we expected neurons exposed to instantaneous checkerboard inversions (flashes) to become more excitable, but surprisingly, they were on average less spiky, and generated 0.4$\pm$0.4 spikes, across all types of dynamic clamp inputs, compared to 0.9$\pm$1.0 spikes in control (Figure 1D, left column; F(1,677)=30.4, p=5e-8, n cells=28, 29). (Here and below, F-values are reported for a multivariate fixed effects analysis of variance with selected interactions, where cell id is used as a repeated measures factor; see Methods for a detailed description). Neurons exposed to looming transitions also spiked less than control neurons (0.6$\pm$0.4 spikes; F(1,629)=9.6; p=0.002; n=28, 25), but more than those exposed to instantaneous "flashes" (F(1,641)=3.7; p=4e-6; n=29, 25). This was likewise unexpected, as compared to flashes, looming stimuli are known to elicit stronger tectal responses \citep{khakhalin2014,khakhalin2019graph}, yet their effect on neuronal excitability was weaker, seemingly in contradiction with the principle of homeostatic plasticity \citep{pratt2007,turrigiano2007}.

@@

We then examined whether different types of sensory activation would differently reshape temporal intrinsic tuning in tectal neurons. As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (instantaneous flashes; \citealt{khakhalin2014}) would selectively suppress spiking in response to faster synaptic inputs. As an alternative to this hypothesis, however, one could argue that exposure to shorter stimuli could make neurons better adjusted to working with fast synaptic currents \citep{stemmler1999information}, similar to changes described for synaptic plasticity \citep{aizenman2007} and recurrent activity in the tectum \citep{pratt2007, shen2011}.

We found that compared to control, neurons exposed to flashes and looming stimuli had different response curves when tested across input conductances of different lengths (F(1, 677)=25.1; p=7e-7; and F(1,629)=12.0; p=6e-4 respectively). In control neurons, longer inputs (500 and 1000 ms) typically evoked stronger spiking than shorter inputs (100 and 200 ms), but neurons from stimulated animals had flatter tuning curves, with a plateau, or even a decrease in spike output for longer conductance injections (Figure 1E). In other words, while control neurons preferred longer inputs, stimulated neurons developed a preference for shorter synaptic inputs, and this change was more pronounced in neurons exposed to flashes than in those exposed to looming stimuli (F(1,641)=7.5; p=0.006). This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, followed two different kinds of logic. The overall excitability was homeostatic, as neurons became less excitable in response to stronger stimulation. The temporal retuning however can be better described as "adaptive", as neurons exposed to shorter stimuli (flashes) became relatively \textit{more} responsive to shorter stimuli, and less responsive to longer stimuli, making them equipped to process faster, synchronous patterns of activation \citep{stemmler1999information, fontaine2014threshold}.

@@

To test this question, we exposed tadpoles to four hours of behaviorally salient, startle-inducing “click” sounds \citep{james2015,truszkowski2017}, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to these sounds (group “Sound”) did not lead to significant changes in either average number of spikes (0.8$\pm$0.7 spikes; F(1,689)=1.7; p=0.2; n=28, 30), amplitude transfer function (F(1,689)=2.0; p=0.2), or temporal tuning curve (F(1,689)=2.0; p=0.2). This may suggest that acoustic stimuli did not activate tectal circuits strongly enough during conditioning, despite being more behaviorally salient (at the onset of stimulation, acoustic clicks evoked startle responses in about 50-80\% of cases, compared to 5-10\% for checkerboard inversions \citep{james2015,truszkowski2017}). This was not necessarily surprising, as mechanosensory and visual inputs have different cellular targets in the tectum \citep{pratt2009trigeminal, felch2016, truszkowski2017}, potentially leading to different recruitment of tectal excitatory and inhibitory circuits. Acoustic stimuli may also be inherently weaker than visual stimuli in triggering plasticity effects in tectal neurons, as they arrive at different compartments within the dendritic tree \citep{hiramoto2009, deeg2009}, which may shape their influence on neuronal plasticity \citep{richards2019dendritic}.

@@

To better match and compare temporal properties of synaptic inputs to those of intrinsic tuning, we calculated average "synaptic current duration" for each cell as a temporal “center of mass” of currents within the first 700 ms after optic chiasm stimulation (see Methods). Neurons with different contribution of early and late synaptic responses naturally had different synaptic current duration: cells with strong monosynaptic inputs had shorter currents, while polysynaptic activity made synaptic currents longer (Figure 3C; p=2e-16, r=$-$0.78, n=168). The synaptic current duration was different across treatment groups (Figure 3D; F(5,163)=6.3, p=2e-5). Cells in Flash, Sync, and Async groups all received shorter synaptic inputs than Control cells (Tukey p$<$0.05, mean duration of 267$\pm$36, 253$\pm$39, 268$\pm$24, and 304$\pm$40 ms, respectively). This shows that prolonged sensory activation with short, frequent stimuli reshaped synaptic transmission in the tectum, making it faster through selective potentiation of visual inputs from the eye, compared to recurrent inputs within the tectum, that remained unchanged. In this study, we did not probe changes in acoustic and mechanosensory projections from the hindbrain.

@@

As our technical resources were rather limited, in this study we don't explicitly address the mechanisms of this newly discovered intrinsic temporal tuning. We can, however, offer two working hypotheses that may explain these results. As tadpole tectal cells don't express "true" resonance currents, such as h-currents \citep{ciarleglio2015}, most temporal tuning effects we observed seem to be due to differences in ionic current inactivation in different cells \citep{azouz2000threshold, fontaine2014threshold}. One obvious way of tuning ionic channel inactivation would be for every cell to modulate sodium or transient potassium currents via channel phosphorylation, or by changing the expression of different channel variants, to shift their inactivation dynamics \citep{frank2003nachannels,goldwyn2018a_current}. To test this hypothesis, one would need to induce changes in temporal tuning similarly to how we did in this paper, pharmacologically isolate different ionic currents, and directly measure their inactivation dynamics \citep{zbili2019axonnav}.

@@

Our findings lead to several verifiable predictions. As rapid inactivation of spiking in tectal neurons plays a role in collision detection \citep{khakhalin2014,jang2016}, a change in temporal tuning should affect collision detection dynamics, which can be verified experimentally. More specifically, we predict that a visual stimulation that retunes neurons to faster stimuli (Flash) would make tadpoles selectively less responsive to slow collisions, and increase the latency of collision avoidance due to non-linear dynamics of retinal activation for realistic looming stimuli that are slow in the beginning and fast towards the end \citep{jang2016}. These changes in intrinsic temporal tuning would also reshape the connectivity of tectal networks, as fast-inactivating cells would not support short recurrent loops within the network, thus promoting long-ranged polysynaptic connectivity \citep{fiete2010stdp,clopath2010stdp}. Finally, based on the multisensory phenomena reported in this paper, we predict that even though multisensory stimulation tends to increase tectal responses in vitro \citep{felch2016,truszkowski2017}, it may reduce peak activation in vivo.
